---
title: "data_analysis"
output: html_notebook
---


#### import dataset and data cleaning
```{r}
library(dplyr)
library(haven)
library(ggplot2)
dataset <- read_stata("data/Regression-and-Event-Study-Analyses.dta")
```


#### Descriptive statistics
```{r descriptive}
#demographics
#all
dataset %>%
  select(percent_female,percent_black,percent_nonwhite,percent_white,percent_other_nonwhite,percent_foreign,percent_age_less15,percent_age_15to44,percent_age_45plus) %>%
  summarise_all(list(mean = ~mean(.), sd = ~sd(.)))
#treatment group
dataset %>%
  filter(filter==1) %>%
  summarise(n=n())
dataset %>%
  filter(filter==1) %>%
  select(percent_female,percent_black,percent_nonwhite,percent_white,percent_other_nonwhite,percent_foreign,percent_age_less15,percent_age_15to44,percent_age_45plus) %>%
  summarise_all(list(mean = ~mean(.), sd = ~sd(.))) 
#control group
dataset %>%
  filter(filter<1) %>%
  summarise(n=n())
dataset %>%
  filter(filter<1) %>%
  select(percent_female,percent_black,percent_nonwhite,percent_white,percent_other_nonwhite,percent_foreign,percent_age_less15,percent_age_15to44,percent_age_45plus) %>%
  summarise_all(list(mean = ~mean(.), sd = ~sd(.)))

#Outcomes: filtration
#all
dataset %>%
  select(mortality_rate,mortality_under1_rate,typhoid_rate,diarrhea_rate) %>%
  summarise_all(list(mean = ~mean(.,na.rm = T), sd = ~sd(.,na.rm = T)))
#treatment
dataset %>%
   filter(filter==1) %>%
  select(mortality_rate,mortality_under1_rate,typhoid_rate,diarrhea_rate) %>%
  summarise_all(list(mean = ~mean(.,na.rm = T), sd = ~sd(.,na.rm = T)))
#control
dataset %>%
  filter(filter<1) %>%
  select(mortality_rate,mortality_under1_rate,typhoid_rate,diarrhea_rate) %>%
  summarise_all(list(mean = ~mean(.,na.rm = T), sd = ~sd(.,na.rm = T)))

test <- dataset %>%
  filter(filter<1 & filter>0)

#Outcomes: chlorination
#treatment
dataset %>%
   filter(chlorination==1) %>%
  select(mortality_rate,mortality_under1_rate,typhoid_rate,diarrhea_rate) %>%
  summarise_all(list(mean = ~mean(.,na.rm = T), sd = ~sd(.,na.rm = T)))
#control
dataset %>%
  filter(chlorination<1) %>%
  select(mortality_rate,mortality_under1_rate,typhoid_rate,diarrhea_rate) %>%
  summarise_all(list(mean = ~mean(.,na.rm = T), sd = ~sd(.,na.rm = T)))

#Outcomes: clean water project
#treatment
dataset %>%
   filter(water_project==1) %>%
  select(mortality_rate,mortality_under1_rate,typhoid_rate,diarrhea_rate) %>%
  summarise_all(list(mean = ~mean(.,na.rm = T), sd = ~sd(.,na.rm = T)))
#control
dataset %>%
  filter(water_project<1) %>%
  select(mortality_rate,mortality_under1_rate,typhoid_rate,diarrhea_rate) %>%
  summarise_all(list(mean = ~mean(.,na.rm = T), sd = ~sd(.,na.rm = T)))
```


### RD Regression

Model specification:

$ln(mortality_{t}) = \beta_0 + \boldsymbol{X}_{t} \beta_1 + \boldsymbol{Z}_t\beta_2+ \beta_3 t + \beta_4 I(t > t^*) + \epsilon_{t}$

$ln(mortality_{t}) = \beta_0 + \boldsymbol{X}_{t} \beta_1 + \boldsymbol{Z}_t\beta_2 + \beta_3 t + \beta_4 I(t > t^*) + \beta_5 I(t > t^*)* t  + \epsilon_{t}$

$ln(mortality_{t}) = \beta_0 + \boldsymbol{X}_{t} \beta_1 + \beta_2 t + \beta_3 I(t > t^*) + \epsilon_{t}$

$ln(mortality_{t}) = \beta_0 + \boldsymbol{X}_{t} \beta_1 + \beta_2 t + \beta_3 I(t > t^*) + \beta_4 I(t > t^*)* t  + \epsilon_{t}$

Data cleaning
```{r data cleaning}
dat <- dataset %>% 
  select(year,
         city,
         city_id,
         percent_female,
         percent_nonwhite,
         percent_foreign,
         percent_age_less15,
         percent_age_15to44,
         percent_age_45plus, 
         ln_mortality_rate,
         ln_mortality_under1_rate,
         ln_typhoid_rate,
         ln_diarrhea_rate,
         mortality_rate,
         mortality_under1_rate,
         typhoid_rate,
         diarrhea_rate,
         filter, chlorination, water_project) %>% 
    mutate(filter = ifelse(filter == 1, 1, 0),
         chlorination = ifelse(chlorination == 1, 1, 0),
         water_project = ifelse(water_project == 1, 1, 0))

start_year <- dat %>% 
  arrange(city, by_group = year) %>% 
  mutate(filter1 = ifelse(filter == 0, 10000, year),
          ch1 = ifelse(chlorination == 0, 10000, year),
         water1 = ifelse(water_project == 0, 10000, year)) %>% 
  group_by(city) %>% 
  summarise(filter_start = min(filter1),
            ch_start = min(ch1),
            water_start = min(water1)) %>% 
  mutate(filter_start = ifelse(filter_start == 10000, 1940, filter_start),
         ch_start = ifelse(ch_start == 10000, 1940, ch_start),
         water_start = ifelse(water_start == 10000, 1940, water_start))

dat <- left_join(dat, start_year) %>% 
  mutate(filter_t = year - filter_start,
         ch_t = year - ch_start,
         water_t = year - water_start,
         filter_knot = ifelse(filter_t <=0, 0, 1),
         ch_knot = ifelse(ch_t <= 0, 0, 1),
         water_knot = ifelse(water_t <= 0, 0, 1)) 


```

## treatment effect of water filteration
```{r}
summary(glm(ln_mortality_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + filter_t + filter_knot, data = dat))

summary(glm(ln_mortality_under1_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + filter_t + filter_knot, data = dat))

summary(glm(ln_typhoid_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + filter_t + filter_knot, data = dat))

summary(glm(ln_diarrhea_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + filter_t + filter_knot, data = dat))
```

## treatment effect of chlorination
```{r}
summary(glm(ln_mortality_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + ch_t + ch_knot, data = dat))

summary(glm(ln_mortality_under1_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + ch_t + ch_knot, data = dat))

summary(glm(ln_typhoid_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + ch_t + ch_knot, data = dat))

summary(glm(ln_diarrhea_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + ch_t + ch_knot, data = dat))
```

### clean water project
```{r}
summary(glm(ln_mortality_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + water_t + water_knot, data = dat))

summary(glm(ln_mortality_under1_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + water_t + water_knot, data = dat))

summary(glm(ln_typhoid_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + water_t + water_knot, data = dat))

summary(glm(ln_diarrhea_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + water_t + water_knot, data = dat))

```


### RD model 2

```{r}
summary(glm(ln_mortality_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + filter_t + filter_knot + filter_knot*filter_t, data = dat))

summary(glm(ln_mortality_under1_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + filter_t + filter_knot + filter_knot*filter_t, data = dat))

summary(glm(ln_typhoid_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + filter_t + filter_knot*filter_t, data = dat))

summary(glm(ln_diarrhea_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + filter_t + filter_knot + filter_knot*filter_t, data = dat))
```

## chlorination
```{r}
summary(glm(ln_mortality_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + ch_t + ch_knot*ch_t, data = dat))

summary(glm(ln_mortality_under1_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + ch_t + ch_knot*ch_t, data = dat))

summary(glm(ln_typhoid_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + ch_t + ch_knot*ch_t, data = dat))

summary(glm(ln_diarrhea_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + ch_t + ch_knot*ch_t, data = dat))
```

## plot observed data with fitted line
```{r}
#all cause mortality
dat.pred.nocov <- 
  dat %>% 
  select(ch_t,ch_knot) %>%
  distinct()

#unadjusted model
mod.mort.nocov <- glm(ln_mortality_rate ~ ch_knot*ch_t, data = dat)
#generate predicted values based on the model in order to create a plot
y.mort.nocov <- predict(mod.mort.nocov, newdata = dat.pred.nocov, type = "response")
#to create the counter-factual scenario we create a data frame as if the intervention was never being implemented
dat.pred.nocov.ct <- 
  dat %>% 
  select(ch_t,ch_knot) %>%
  filter(ch_knot==1) %>%
  mutate(ch_knot=0) %>%
  distinct()
#generate fitted value for counter-factual scenario
y.mort.nocov.ct <- predict(mod.mort.nocov, newdata = dat.pred.nocov.ct, type = "response")

#plot
fig_mort <- ggplot() +
  geom_point(aes(x = dat$ch_t, y = dat$ln_mortality_rate),color="darkgrey") +
  geom_line(aes(x = dat.pred.nocov$ch_t, y = y.mort.nocov),color="red") + # fitted line
  geom_line(aes(x = dat.pred.nocov.ct$ch_t, y = y.mort.nocov.ct),color="red",linetype="dashed") +
  geom_vline(xintercept = 0, color = "blue") +
  labs(x = "year relative to start of treatment", y = "log all-cause mortality")+
  theme(legend.position = "none") +
  theme_bw()

#adjusted model
dat.pred <- dat %>% select(
                           percent_female,
                           percent_nonwhite,
                           percent_foreign,
                           percent_age_less15,
                           percent_age_15to44,
                           percent_age_45plus,
                           ch_t, 
                           ch_knot)

mod.mort <- glm(ln_mortality_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + ch_t + ch_knot*ch_t, data = dat)

y.mort <- predict(mod.mort, newdata = dat.pred)

ggplot() +
  geom_point(aes(x = dat$ch_t, y = dat$ln_mortality_rate)) +
  geom_point(aes(x = dat$ch_t, y = y.mort, color = as.factor(dat$ch_knot + 1))) +
  geom_vline(xintercept = 0, color = "blue") +
  labs(x = "year relative to start of treatment", y = "log all-cause mortality")+
  theme(legend.position = "none")


#infant mortality
#unadjusted model
mod.in.nocov <- glm(ln_mortality_under1_rate ~ ch_knot*ch_t, data = dat)
#generate predicted values based on the model in order to create a plot
y.in.nocov <- predict(mod.in.nocov, newdata = dat.pred.nocov, type = "response")
#generate fitted value for counter-factual scenario
y.in.nocov.ct <- predict(mod.in.nocov, newdata = dat.pred.nocov.ct, type = "response")

#plot
fig_in <- ggplot() +
  geom_point(aes(x = dat$ch_t, y = dat$ln_mortality_under1_rate),color="darkgrey") +
  geom_line(aes(x = dat.pred.nocov$ch_t, y = y.in.nocov),color="red") + # fitted line
  geom_line(aes(x = dat.pred.nocov.ct$ch_t, y = y.in.nocov.ct),color="red",linetype="dashed") +
  geom_vline(xintercept = 0, color = "blue") +
  labs(x = "year relative to start of treatment", y = "log infant mortality")+
  theme(legend.position = "none") +
  theme_bw()

#adjusted model
mod.in <- glm(ln_mortality_under1_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + ch_t + ch_knot*ch_t, data = dat)

y.in <- predict(mod.in, newdata = dat.pred)

ggplot() +
  geom_point(aes(x = dat$ch_t, y = dat$ln_mortality_under1_rate) )+
  geom_point(aes(x = dat$ch_t, y = y.in, color = as.factor(dat$ch_knot + 1))) +
  geom_vline(xintercept = 0, color = "blue") +
  labs(x = "year relative to start of treatment", y = "log infant mortality")+
  theme(legend.position = "none")


#typhoid mortality
#unadjusted model
mod.ty.nocov <- glm(ln_typhoid_rate ~ ch_knot*ch_t, data = dat)
#generate predicted values based on the model in order to create a plot
y.ty.nocov <- predict(mod.ty.nocov, newdata = dat.pred.nocov, type = "response")
#generate fitted value for counter-factual scenario
y.ty.nocov.ct <- predict(mod.ty.nocov, newdata = dat.pred.nocov.ct, type = "response")

#plot
fig_ty <- ggplot() +
  geom_point(aes(x = dat$ch_t, y = dat$ln_typhoid_rate),color="darkgrey") +
  geom_line(aes(x = dat.pred.nocov$ch_t, y = y.ty.nocov),color="red") + # fitted line
  geom_line(aes(x = dat.pred.nocov.ct$ch_t, y = y.ty.nocov.ct),color="red",linetype="dashed") +
  geom_vline(xintercept = 0, color = "blue") +
  labs(x = "year relative to start of treatment", y = "log typhoid mortality")+
  theme(legend.position = "none") +
  theme_bw()

#adjusted model
mod.ty <- glm(ln_typhoid_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + ch_t + ch_knot*ch_t, data = dat)

y.ty <- predict(mod.ty, newdata = dat.pred)

ggplot() +
  geom_point(aes(x = dat$ch_t, y = dat$ln_typhoid_rate) )+
  geom_point(aes(x = dat$ch_t, y = y.ty, color = as.factor(dat$ch_knot + 1))) +
  geom_vline(xintercept = 0, color = "blue") +
  labs(x = "year relative to start of treatment", y = "log typhoid mortality")+
  theme(legend.position = "none")


#diarrhea mortality
#unadjusted model
mod.di.nocov <- glm(ln_diarrhea_rate ~ ch_knot*ch_t, data = dat)
#generate predicted values based on the model in order to create a plot
y.di.nocov <- predict(mod.di.nocov, newdata = dat.pred.nocov, type = "response")
#generate fitted value for counter-factual scenario
y.di.nocov.ct <- predict(mod.di.nocov, newdata = dat.pred.nocov.ct, type = "response")

#plot
fig_di <- ggplot() +
  geom_point(aes(x = dat$ch_t, y = dat$ln_diarrhea_rate),color="darkgrey") +
  geom_line(aes(x = dat.pred.nocov$ch_t, y = y.di.nocov),color="red") + # fitted line
  geom_line(aes(x = dat.pred.nocov.ct$ch_t, y = y.di.nocov.ct),color="red",linetype="dashed") +
  geom_vline(xintercept = 0, color = "blue") +
  labs(x = "year relative to start of treatment", y = "log diarrhea mortality")+
  theme(legend.position = "none") +
  theme_bw()

#adjusted model
mod.di <- glm(ln_diarrhea_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + ch_t + ch_knot*ch_t, data = dat)


y.di <- predict(mod.di, newdata = dat.pred)

ggplot() +
  geom_point(aes(x = dat$ch_t, y = dat$ln_diarrhea_rate) )+
  geom_point(aes(x = dat$ch_t, y = y.di, color = as.factor(dat$ch_knot + 1))) +
  geom_vline(xintercept = 0, color = "blue") +
  labs(x = "year relative to start of treatment", y = "log diarrhea mortality")+
  theme(legend.position = "none")

#combine the four unadjusted plots to one and save the figure
#required package
#if(!require(devtools)) install.packages("devtools")
#devtools::install_github("kassambara/ggpubr")
library(ggpubr)
# 1. Open jpeg file
jpeg("unadj_ITS_chl.jpg", width = 12, height = 8, units = 'in', res = 500)
# 2. Create the plot
ggarrange(fig_mort, fig_in, fig_ty, fig_di, 
          labels = c("A", "B", "C", "D"),
          ncol = 2, nrow = 2)
# 3. Close the file
dev.off()
```




## water project
```{r}
summary(glm(ln_mortality_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + water_t + water_knot*water_t, data = dat))

summary(glm(ln_mortality_under1_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + water_t + water_knot*water_t, data = dat))

summary(glm(ln_typhoid_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + water_t + water_knot*water_t, data = dat))

summary(glm(ln_diarrhea_rate ~ percent_female + percent_nonwhite + percent_foreign + percent_age_less15 + percent_age_15to44 + percent_age_45plus + water_t + water_knot*water_t, data = dat))
```

Code from Interrupted time series regression for the evaluation of public health interventions: a tutorial by James Lopez Bernal, Steven Cummins, Antonio Gasparrini
```{r}
# Install packages required for the analysis (uncomment if needed)
#install.packages("lmtest") ; install.packages("Epi")
#install.packages("tsModel"); install.packages("vcd")

# load the packages
library(foreign)
library(tsModel)
library("lmtest")
library("Epi")
library("splines")
library("vcd")

## Scatter plot
# start the plot, excluding the points and the x-axis
plot(dat$mortality_rate,type="n",ylim=c(600,2800),xlab="Year", ylab="Std rate x 100,000",
  bty="l")
# shade the post intervention period grey
rect(36,0,60,300,col=grey(0.9),border=F)
# plot the observed rate for pre-intervention period
points(dat$mortality_rate,cex=0.7)
#specify the x-axis (i.e. time units)
axis(1,at=0:25*12,labels=F)
axis(1,at=0:40*12,tick=F)
# add a title
title("Sicily, 2002-2006")

```


```{r}
#example
# read data from csv file
data <- read.csv("sicily.csv")
head(data)
View(data)

#Poisson with the standardised population as an offset
model1 <- glm(aces ~ offset(log(stdpop)) + smokban + time, family=poisson, data)
summary(model1)
summary(model1)$dispersion
round(ci.lin(model1,Exp=T),3)

# create a new dataframe with 0.1 time units to improve the graph
datanew <- data.frame(stdpop=mean(data$stdpop),smokban=rep(c(0,1),c(360,240)),
  time= 1:600/10,month=rep(1:120/10,5))

# We generate predicted values based on the model in order to create a plot
pred1 <- predict(model1,type="response",datanew)/mean(data$stdpop)*10^5

#This can then be plotted along with a scatter graph (see above)
plot(data$rate,type="n",ylim=c(0,300),xlab="Year",ylab="Std rate x 10,000",
  bty="l",xaxt="n")
rect(36,0,60,300,col=grey(0.9),border=F)
points(data$rate,cex=0.7)
axis(1,at=0:5*12,labels=F)
axis(1,at=0:4*12+6,tick=F,labels=2002:2006)
lines((1:600/10),pred1,col=2)
title("Sicily, 2002-2006")

# to plot the counterfactual scenario we create a data frame as if smokban
#   (the intervention) was never being implemented
datanew <- data.frame(stdpop=mean(data$stdpop),smokban=0,time=1:600/10,
  month=rep(1:120/10,5))

# generate predictions under the counterfactual scenario and add it to the plot
pred1b <- predict(model1,datanew,type="response")/mean(data$stdpop)*10^5
lines(datanew$time,pred1b,col=2,lty=2)

# return the data frame to the scenario including the intervention
datanew <- data.frame(stdpop=mean(data$stdpop),smokban=rep(c(0,1),c(360,240)),
  time= 1:600/10,month=rep(1:120/10,5))

```


